---
title: "Assignment 6"
subtitle: "STAT 201"
author: "Irene Li"

format:
  html:
    toc: true
    embed-resources: true

execute:
  message: false
  warning: false
---

## Instructions {-}

1. **Write your name** on the assignment

2. Write your code in the *Code* cells of the **template provided** to write solutions for the assignment. **Do not open a new qmd**, and work from scratch. Ensure that the solution is written neatly enough to understand and grade.

3. You may talk to a friend, discuss the questions and potential directions for solving them. However, you need to write your own solutions and code separately, and not as a group activity. Do not use AI to solve the problems.

4. There are point deductions if your code is not clean and organized.

5. Only print what is asked for. Highly inefficient code will be deducted for because it demonstrates lack of understanding.

6. Render as an HTML and **submit BOTH the .qmd and HTML file**.

The grader cannot grade what they cannot see! Failure to render your document after completing all questions will lead to deductions.


## Question 1 (3 points)

### Part a

Create a vector called `bears` which contains the values `20, 17, 10, 28, 13, 30, 13, 17, 16, 26, 12, 28, 17, 27`.

Create a vector called `opponent` which contains the values `27, 41, 31, 20, 19, 12, 30, 24, 13, 31, 10, 13, 20, 16`.

Note: each element represents the amount of points scored in a particular football game.

No need to print any output here.

```{r}
bears <- c(20, 17, 10, 28, 13, 30, 13, 17, 16, 26, 12, 28, 17, 27)
opponent <- c(27, 41, 31, 20, 19, 12, 30, 24, 13, 31, 10, 13, 20, 16)
```


### Part b

Insert the value `40` as the 5th element in `bears`.

Insert the value `38` as the 1st element in `opponent`.

Print both vectors to prove you were successful.

```{r}
bears <- append(bears, 40, after=4)
opponent <- append(opponent, 38, after=0)

cat("Bears scores: ", bears, "\n")
cat("Opponent scores: ", opponent)
```


### Part c

Determine the largest win spread for the Bears (ie: spread is the difference between bear's score and opponent's score) and print the following statement:

"The largest win spread is {}, when the Bears scored {} points and the Opponent scored {} points."

Note: this needs to still run correctly even if the elements changed or were re-ordered.

```{r}
spread <- bears - opponent

#find index of largest spread
index_large <- 0
max_num <- max(spread)
index_large <- which(spread == max_num)

#get numbers from index
bears_score <- bears[index_large]
opponent_score <- opponent[index_large]
win_spread <- spread[index_large]

cat("The largest win spread is", win_spread, "when the Bears scored ", bears_score, " points and the Opponent scored ", opponent_score, " points")


```


## Question 2 (7 points)

### Part a

Create a matrix `game_score` by utilizing the vectors `bears` and `opponent` from Question 1 where the first column represents `bears` and the second column represents `opponent`. 

Name the columns `bears` and `opponent` respectively.

Print `game_score` to prove this was successful.

```{r}
game_score <- cbind(bears, opponent)
colnames(game_score) <-c("Bears", "Opponent")
print(game_score)
```


### Part b

The `game_score` is missing some games. Add on a matrix that contains bears scores of `37, 9` and opponent scores of `17, 17`, respectively. Print the new `game_score` matrix.

```{r}
missing_scores <- matrix(c(37,9,17,17), ncol=2)
game_score <- rbind(game_score, missing_scores)
print(game_score)
```


### Part c

Iterate through `game_score` and print "Win" if the bears won and "Loss" if the bears lost.

```{r}
#iterate through rows with i
for (i in 1:nrow(game_score)){
  if (game_score[i,"Bears"] > game_score[i, "Opponent"]){
    print("Win")
  } else {
    print("Loss")
  }
}

```


### Part d

Use `game_score` and an **inequality** to check if the bears won each game. Store this as a vector called `win` and print the vector `win`.

ie: Since 20 is less than 38 the first element of `win` should be FALSE.

```{r}
#create new vector to store wins and losses
win <- c()
for (i in 1:nrow(game_score)){
    win <- append(win, as.logical(game_score[i,"Bears"] > game_score[i, "Opponent"])) 
}
print(win)
```


### Part e

Add the vector `win` as a third column to `game_score`. Print `game_score`.

```{r}
game_score <- cbind(game_score, win)
print(game_score) #matrices can only store 1 data type so it converted my TRUE/FALSE to 0s and 1s
```


### Part f

How many games did the Bears lose before their first win (games are in order)? Many ways to accomplish this BUT you must utilize `game_score` to do so. 

Clarification: do not directly reference any of the individual vectors from previous question parts, this must use the matrix `game_score`. 

The code must be capable of running correctly even if the values changed.

```{r}
lose_count <- 0
for (i in 1:nrow(game_score)){
  if (game_score[i,"win"] == 0){
    lose_count <- lose_count + 1
  }
  else {
      break
    }
  }

cat("the bears lost", lose_count ,"games before their first win")
```


## Question 3 (2 points)

Write a function called `number_of_factors` that takes an integer and returns how many factors the number has. Run your function with `8` to check if it works.

```{r}
number_of_factors <- function(num){
  factors = 0
  for (i in 1:num){
    if ((num %% i) == 0){
      factors <- factors + 1
    }
  }
  return(factors)
}
num <- 8
factors <- number_of_factors(num)
cat(num, "has", factors, "factors")
```


## Question 4 (3 points)

Write a function that calculates the area of a rectangle. The function should have an input for length (in inches) and width (in inches). Set a default value for both length and width to be 48.

The function should have 2 outputs: the area in terms of square inches and the area in terms of square feet. There are 12 inches in 1 foot.

Run your function with a length of 102 and width of 60 and store the output as a variable. Then print the following:

"The area in square inches is {}."

"The area in square feet is {}."

```{r}
#writing function
area_rec <- function(length=48, width=48){
  #get 
  length_in <- length
  width_in <- width
  length_ft <- length_in/12
  width_ft <- width_in/12
  #
  area_in <- length_in * width_in
  area_ft <- length_ft * width_ft
  vec <- c(area_in, area_ft)
  return(vec)
}

#printing the results
result <- area_rec(length = 102, width = 60)
cat(paste0("The area in square inches is ", result[[1]], ".\n"))
cat(paste0("The area in square feet is ", result[[2]], "."))
```


## Question 5 (5 points)

Write a function that takes a single string vector as an input. Have the function return only the word (or words) with the most letters. 

If there are multiple words with the most letters, then return all words with the most letters **as a vector**.

Clarification: This is NOT returning a list that contains each word individually. It is returning a vector containing all words with the most letters.

Test the function using the vector `fruits` created below.

```{r}
fruits <- c("apple", "orange", "banana", 
            "pear", "grapes", "kiwi")
```


```{r}
longest_word <- function(vector){
  lengths <- nchar(vector) #make a vector of string lengths
  max_length <- max(lengths) #get max length
  result <- vector[lengths == max_length] #returns elements in vector where lengths == max_length is TRUE
  return(result)
}
print(longest_word(fruits))
```


## Question 6 (4 points)

Write a function that checks if a value is in a vector.
The function should take both a value and vector as inputs and return either TRUE or FALSE.
  
Run your function using the variables `value` and `check` created below.
  
```{r}
value <- 1
check <- seq(1, 15, 3)

#just showing myself the vector ignore this
print(check)
```

After running the function, depending on what the function returns, print either:

  - "{} is in the vector."
  - "{} is not in the vector.
  
```{r}
#writing functions
checking <- function(value, vector){
  return(value %in% vector)
}
#trying function
tf <- checking(value, check)
if (tf == TRUE){
  cat(value, "is in the vector.")
} else {
  cat(value, "is not in the vector.")
}
```


## Question 7 (6 points)

Below is a matrix of random numbers.

```{r}
# Don't change this! A seed keeps results consistent.
set.seed(123)
x <- matrix(sample(-50:50, 20), nrow = 5)
print(x)
```

### Part a   

Write a function that takes a matrix as an input, converts all negative numbers to NA, and returns the new matrix. Test your function using the matrix `x` that was created for you.

```{r}

to_na <- function(matrix){
  matrix[matrix < 0] <- NA #replaces all negative numbers with NA
  return(matrix)
}
print(to_na(x))
```

### Part b

Write a function that takes a matrix as an input and converts all even numbers to 1 and odd numbers to 0. The function should return the new matrix. Test your function using the matrix `x` that was created for you.

```{r}
even_odd <- function(matrix){
  num_row <- nrow(matrix)
  num_col <- ncol(matrix)
  for (i in 1:num_row){
    for (j in 1:num_col){
      if (matrix[i,j] %% 2 == 0){ #even
        matrix[i,j] <- 1
      } else { #odd
        matrix[i,j] <- 0
      }
    }
  }
    return(matrix)
}
print(even_odd(x))
```

